\documentclass[journal]{IEEEtran}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{url}
\usepackage[hidelinks,breaklinks=true,bookmarks=false]{hyperref}

% The manuscript body is split into per-section files under sections/ and pulled
% in with \input below, so the prose can be edited independently of the preamble.

\begin{document}

\title{MelanomaNet: Explainable Deep Learning for\\ Skin Lesion Classification}

\author{Sukhrobbek~Ilyosbekov%
    \thanks{S. Ilyosbekov is with Northeastern University, Boston, MA, USA
        (e-mail: ilyosbekov.s@northeastern.edu).}%
    \thanks{Manuscript submitted for review. A preprint of this work is
        available on arXiv.}%
}

\markboth{}{Ilyosbekov: Explainable Deep Learning for Skin Lesion Classification}

\maketitle

\begin{abstract}
    Deep neural networks classify dermoscopic skin lesions with an accuracy that
    approaches that of experienced dermatologists, but they are seldom used in
    routine practice because they rarely expose the basis for a prediction. We
    describe MelanomaNet, a skin-lesion classifier that pairs an EfficientNetV2-M
    backbone with four post-hoc interpretability components: GradCAM++ attention
    maps, automated extraction of the clinical ABCDE criteria, concept activation
    vectors trained with a stochastic-gradient classifier, and Monte Carlo dropout
    uncertainty estimates. The system is trained and evaluated on the ISIC 2019
    dermoscopy dataset (25,331 images; eight populated diagnostic classes, with a
    ninth ``unknown'' category that contains no labelled images). On the held-out
    test set the classifier reaches 81.7\% balanced accuracy and a macro-averaged
    F1 of 0.80 (85.6\% overall accuracy, weighted F1 0.856), the drop between the
    two reflecting the marked class imbalance. The uncertainty component separates
    epistemic from aleatoric contributions and flags low-confidence cases for
    review, and an alignment analysis relates the attention maps to the segmented
    lesion and its border. We report these results not as a claim of clinical
    readiness but as evidence that competitive accuracy and inspectable predictions
    can coexist in a single model. Source code and trained weights are available at
    \url{https://github.com/suxrobgm/explainable-melanoma}.
\end{abstract}

\begin{IEEEkeywords}
    Explainable AI, melanoma, dermoscopy, skin lesion classification,
    uncertainty quantification, concept activation vectors, deep learning.
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

\input{sections/introduction}
\input{sections/related_work}
\input{sections/methods}
\input{sections/experiments}
\input{sections/discussion}
\input{sections/conclusion}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
